Strength, mating success, and immune and dietary costs in a population sample of US women and men: a registered report

### 2024-09-14 paper.qmd
Minor edits. Updated index.html

## Post Hoc Analyses
We were made aware that Dual-Energy X-ray Absorptiometry (DXA) data were published for these series after our initial download of the data for this study. These data provide alternatives to grip strength for assessing muscle mass, and include the lean masses (i.e., fat-free mass, optionally including or excluding bone mineral content) in grams of each arm and leg, the trunk, and total body. Total body lean mass is similar to the bioelectrical impedance analysis fat-free mass variable used by @lassek2009. For reviews of DXA measurements of body composition, see @shepherd2017 and @messina2020. For details of the NHANES DXA measurements, see @DXXG2023. In addition to our grip strength models, we also fit our models exactly as specified in stage 1 except substituting four DXA lean masses for grip strength (all masses excluded bone mineral content). Specifically, we computed arm lean mass as the sum of left and right arm lean masses, and leg lean mass as the sum of left and right leg lean masses. We then fit our models using arm, leg, trunk, and total lean masses, each standardized by sex as described earlier for grip strength.

@@

Unlike grip strength, arm, leg, and total (but not trunk) lean masses were all negatively associated with WBCC. We found that all lean masses were significant, positive predictors of both energy and protein intake controlling for sex. In the expanded models with additional controls, all lean masses remained a significant predictor of energy and protein intake. See @sfig-arms - @sfig-total.

@@

Our registered report had one aim: to test the prediction from the sexual selection hypothesis that the effect of upper body strength on reproductive success is significantly positive for males and significantly less so for females. Our aim was not to test all possible byproduct hypotheses, nor was it possible to do so. We were unable to control for in-utero testosterone exposure, or other conditions such as Polycystic Ovarian Syndrome (PCOS) that might influence the relationship between grip strength and our measures of mating success. Likewise, NHANES does not collect data on sociosexuality. As described in our methods, in industrialized populations like the U.S., widespread access to contraceptives uncouples reproductive success from mating behavior for both men and women. Measures of mating success, including partnered status, number of sexual partners, and age at first sexual intercourse, are used as proxies as they are assumed to have been strongly correlated with reproductive success under ancestral conditions [@perusse1993]. In addition, in confirmatory analyses of mating outcomes (but not immune, energy, or protein outcomes) we removed `{r} sum(designsH$d.design.adults$variables$sex_partners>=100, na.rm = T)` participants with 100 or more lifetime sexual partners on the grounds that these might involve evolutionarily novel patterns of behavior that would not conform to predictions of the sexual selection hypothesis. Ideally, however, analyses would account for the full range of variation in sexual behavior. Finally, all pilot results were obtained after considerable exploratory analysis, with a high risk of overfitting the data, and therefore our confirmatory results were critical to testing our predictions. However, even our confirmatory results were based on generalized linear models, yet there might be important non-linear relationships that would alter interpretations.

@@

Moreover, in addition to formidability, which is important in male-male competition for short and long-term partners, upper body strength indexes productivity. Specifically, strength in forearm and shoulder is critical for effective use of projectile hunting technology such as the bow and arrow [@ertan2003] and javelin [@kim2014]. Indeed, @apicella2014 found that upper body strength predicted both hunting reputation and reproductive success in Hadza men. The similarities in our results between grip strength and arm lean mass as predictors further supports the importance of upper body strength in particular. Therefore, our results could support an evolved female preference for productivity, i.e., the greater ability over human evolution of stronger men to provision offspring through big-game hunting [@kaplan2000; @apicella2014], although an evolved preference for long-term partners who can confer protection from predators or other men is possible too [@willems2017b; @kelly2005].

>>>>

## Post Hoc Analyses
We were made aware that Dual-Energy X-ray Absorptiometry (DXA) data were published for these series after our initial download of the data for this study. DXA data provide alternatives to grip strength for assessing muscle mass, and include the lean masses (i.e., fat-free mass, optionally including or excluding bone mineral content) in grams of each arm and leg, the trunk, and total body. Total body lean mass is similar to the bioelectrical impedance analysis fat-free mass variable used by @lassek2009. For reviews of DXA measurements of body composition, see @shepherd2017 and @messina2020. For details of the NHANES DXA measurements, see @DXXG2023. In addition to our grip strength models, we also fit our models exactly as specified in stage 1 except substituting four DXA lean masses for grip strength (all masses excluded bone mineral content). Specifically, we computed arm lean mass as the sum of left and right arm lean masses, and leg lean mass as the sum of left and right leg lean masses. We then fit our models using arm, leg, trunk, and total lean masses, each standardized by sex as described earlier for grip strength.

@@

Unlike grip strength, arm, leg, and total (but not trunk) lean masses were all negatively associated with WBCC. We found that all lean masses were significant, positive predictors of both energy and protein intake controlling for sex. In the expanded models with additional controls, all lean masses remained significant predictors of energy and protein intake. See @sfig-arms - @sfig-total.

@@

Our registered report had one aim: to test the prediction from the sexual selection hypothesis that the effect of upper body strength on reproductive success is significantly positive for males and significantly less so for females. Our aim was not to test all possible byproduct hypotheses, nor was it possible to do so. We were unable to control for in-utero testosterone exposure, or other conditions such as Polycystic Ovarian Syndrome (PCOS) that might influence the relationship between grip strength and our measures of mating success. Likewise, NHANES does not collect data on sociosexuality. As described in our methods, in industrialized populations like the U.S., widespread access to contraceptives uncouples reproductive success from mating behavior for both men and women. Measures of mating success, including partnered status, number of sexual partners, and age at first sexual intercourse, are used as proxies as they are assumed to have been strongly correlated with reproductive success under ancestral conditions [@perusse1993]. In addition, in confirmatory analyses of mating outcomes (but not immune, energy, or protein outcomes) we removed `{r} sum(designsH$d.design.adults$variables$sex_partners>=100, na.rm = T)` participants with 100 or more lifetime sexual partners (`{r} signif(100*sum(designsH$d.design.adults$variables$sex_partners>=100, na.rm = T)/nrow(designsH$d.design.adults$variables), 3)`%) on the grounds that these might involve evolutionarily novel patterns of behavior that would not conform to predictions of the sexual selection hypothesis. Ideally, however, analyses would account for the full range of variation in sexual behavior. Finally, all pilot results were obtained after considerable exploratory analysis, with a high risk of overfitting the data, and therefore our confirmatory results were critical to testing our predictions. However, even our confirmatory results were based on generalized linear models, but there might be important non-linear relationships that would alter interpretations.

@@

Moreover, in addition to formidability, which is important in male-male competition for short and long-term partners, upper body strength indexes productivity. Specifically, strength in forearm and shoulder is critical for effective use of projectile hunting technology such as the bow and arrow [@ertan2003] and javelin [@kim2014]. Indeed, upper body strength predicted both hunting reputation and reproductive success in Hadza men [@apicella2014]. The similarities in our results between grip strength and arm lean mass as predictors further supports the importance of upper body strength in particular. Therefore, our results could support an evolved female preference for productivity, i.e., the greater ability over human evolution of stronger men to provision offspring through big-game hunting [@kaplan2000; @apicella2014], although an evolved preference for long-term partners who can confer protection from predators or other men is possible too [@willems2017b; @kelly2005].
